---
title: "Defining *spatial closeness* with commuter flow rates"
subtitle: "Leveraging 2011 UK Census data for spatial regression modeling"
author: "Matthew T. West"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  beamer_presentation:
    latex_engine: xelatex
    theme: "metropolis"
    colortheme: "metropolis"
    fonttheme: "metropolis"
csl: british-journal-of-political-science.csl
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, comment=NA, fig.pos='H')
knitr::opts_knit$set(root.dir='~/Documents/brexit_spatial/r_scripts')
```

```{r setup, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
# Load libraries
package_names <- c('readr', 'dplyr', 'ggplot2', 'knitr', 'knitcitations')
for (pkg in package_names) {
    library(pkg, character.only=TRUE, quietly=TRUE, verbose=FALSE)
}

# Root directory where all data in repo is located
data_loc <- '~/Documents/brexit_spatial/data_sets'
# Needed functions to process spatial objects
source('spatial_neighbors.r')
```

## Relationship between separated observations
```{r spdep_obj, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE}
lad.counts <- lad_commuters_2011(data_loc)
# compute percentage flow rates and filter list
nb.pcts <- reduce_neighbors(lad.counts)
# percentages for commuters who work in the district they reside
local_worker_pct <- nb.pcts %>%
    filter(cd.work == cd.reside) %>%
    rename(ladcd = cd.work) %>%
    select(ladcd, r_pct, w_pct)

# construct spatial objects with binary weights via spdep
nb.obj <- define_neighbors(nb.pcts, symm=TRUE, w_type='pct', thresh=0.015)
# Get boundary polygons for english districts
english_bnds <- boundary_polygons(data_loc) %>%
    filter(str_detect(lad16cd, '^E0')) %>%
    select(-c(lad16nm))
```

I work with Bayesian spatial CAR models to investigate the socioeconomic forces underlaying the results of recent \alert{contentious} UK popular referenda at the local authority district (LAD) level. 

 - SOP of defining *closeness* by border contiguity misses two facts
    + Borders are often drawn to mirror physical barriers
    + Modern transit allows people to travel great distances regularly
 - *Proposal* for adjacency weights matrix: $W_{ij}$
    + $w_{ij} =$ % of residents of district $i$ that work in district $j$
    + $\tilde{w}_{ij} =$ % of workers in $j$ that reside in $i$
    + $W_{ij} = \max \left(w_{ij}, \tilde{w}_{ij} \right)$
 - R packages used: `sf, spdep, readr, dplyr, ggplot2`

One can still force $W$ to be *symmetric* and/or have *binary* weights.

## Links between spatially related English LADs
```{r commuter_map, echo=FALSE, eval=TRUE, warning=FALSE, message=FALSE, fig.echo=TRUE, out.width='80%', fig.align="center", fig.cap="The grey polygons represent the 326 LADs in England. The red lines represent the 3286 unique links between districts with $W_{ij} > 0.015$ with the line thickness being related to the entry's weight value."}

map <- commuter_map(nb.obj, english_bnds)
map
```
